The Heavy Photon Search Test Detector

### 2013-11-16 hps-testrun-nim.tex
Added empty acknowledgement section.
Added new t0 figure. Added new pulse shape figure. Added description of t0 resolution stats. Changed discussion of SVT S/N and absolute calibration slightly. Added EoverP for data and MC. Updated the Ecal calibration paragraph slightly. Removed the Ecal EoverP 2D map.

The photon beam was generated in the interaction of the $5.5$ GeV electrons with a gold radiator of 
$10^{-4}~X_0$, located $\approx~9$~m upstream of the PS. The primary beam and scattered 
electrons are deflected away from detectors by the dipole magnet of the photon tagging system. After 
collimation (6.4~mm diameter), the photon beam passes through the aluminum PS pair converter and 
later the HPS system. 
The converter was located $\approx~77$~cm upstream of the first layer of the silicon vertex tracker. Data was 
taken on different converters (empty, $1.8\times 10^{-3}~X_0$ , $4.5\times 10^{-3}~X_0$ and 
$1.6 \times 10^{-2}~X_0$). 
These measurements were repeated with the reverse field setting of the PS dipole magnet.

@@

During the duration of the test run all SVT modules and APV25 chips were configured to their 
nominal operating points~\cite{Jones:1069892} with all sensors reverse-biased at 180~V.  The 
sensors were operated within a temperature range of  $20-24^\circ$C 
throughout the test run. Multiple calibration runs, normalized with lab-based radioactive source tests, 
established a noise level of $\approx$ XXXX-XXXX electrons which was stable across all the SVT modules.

Throughout the duration of the test run, approximately 97\% of the 12,780 SVT channels were found to be 
operating normally. The fraction of dead or noisy channels varied from 2.4\% to 4.7\%.  
Most of these were due to misconfigured readout chips (2--4 misconfigured chips out of 100 ), 
a known noisy half-module and a couple of known noisy readout chips.

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. 
The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svt_daq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/08062012_run1351_samples_vs_amplitude}
    	\caption{The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}

These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}. 
The signal to noise ratio is XXXX.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_mip_small.pdf}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit time 
with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter due to clock 
phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[h]
\begin{center}
{\small
%	\includegraphics[width=7cm]{figures/top_track_time.pdf}
	\includegraphics[width=6cm]{figures/t0_resolution.pdf}
	\caption{
%	Track time distribution (left) and cluster time residual (right). The track time is measured relative to the APV25 clock. The width of the distribution is due to trigger jitter (24~ns jitter in the tracker readout clock, plus 16~ns jitter in the trigger system). 
	The cluster time residual is for a representative sensor relative to the track time.}
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) the extracted $t_0$ resolution is 
2.6~ns. This is somewhat worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate with 
the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

While occupancy was slightly larger than expected, good agreement between data and simulation was 
found after taking into account dead or noisy channels. The hit efficiency, see Fig.~\ref{fig:hit_efficiency}, 
was measured to be above 98\% and fairly uniform across the SVT.
\begin{figure}[h]
\begin{center}
{\small
    	\includegraphics[width=7cmh]{figures/hit_efficiency_vs_layer.pdf}
        \caption{ The hit reconstruction efficiency as a function of detector layer.}
	\label{fig:hit_efficiency}
}
\end{center}
\end{figure}
The spatial resolution of similar microstrip sensors is well established by test beam data, against which the 
charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise (S/N) and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ration above 20, as demonstrated here, 
is relatively constant at approximately 6~$\mu$m for tracks that are close to normal to the sensors as in HPS.

@@

For the vertexing 
performance the foremost difference compared to electron beam running is that the target was 
located approximately 67~cm from our nominal target position; giving almost collinear tracks in the detector. 
This degrades the vertex resolution along the 
beam line compared to that expected in an electron beam with tracks from the nominal target position. 
Furthermore, tails of the vertex distributions are impossible to study with the finite data sample obtained 
in the test run. Nevertheless, useful information can still be 
obtained by studying the vertex distributions. Figure~\ref{fig:vtx_pos} shows the distance of closest 
approach of the momentum vectors extrapolated in the 
upstream direction from our analyzing magnet, taking into account the measured fringe field of the PS magnet. 
 \begin{center}
{\small
	 \begin{figure*}[ht]
	% Plot info:
	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
	\includegraphics[width=5cm]{figures/h_vtx_fr_x_h_vtx_x_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\includegraphics[width=5cm]{figures/h_vtx_fr_y_h_vtx_y_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\includegraphics[width=5cm]{figures/h_vtx_fr_z_h_vtx_z_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\caption{ Vertex position represented by the distance of 
	closest approach of the extrapolated momentum vectors upstream of the PS magnet taking into account 
	the measured fringe field. The overall shift from zero in the x-direction is 
	due to a 30~mrad rotation of the SVT with respect to the beam line.}
	\label{fig:vtx_pos}
\end{figure*}
}
\end{center}
While the tails of the vertex distribution expected in electron beam running is not accessible here the 
fact that the core is relatively well described provides confidence of the description of the amount of 
material and the model used for simulating multiple scattering in the target.  
These are crucial for benchmarking the HPS \Aprime{} physics reach since both the mass and vertex resolution 
that determine the sensitivity are limited by multiple scattering.

@@

The gain of the individual ECal channels has been calibrated using the SVT measurement of track momentum and comparison to Monte Carlo simulation. The integrated pulse of the each channel was converted to energy by first subtracting the pedestal then applying a gain or conversion factor to convert ADC counts to energy in MeV. The pedestals have been measured using special runs where FADCs sampled APD signals every 4~ns and for each trigger a time window of 100 samples have been recorded. Pedestal levels were determined by looking at a part of the 
window before the actual hit. Modules above a threshold are clustered using a simple algorithm described in Sec.~\ref{sec:trigger}. The ECal was operated 
with an effective readout threshold of 73~MeV. The noise level of the ECal modules were measured to be approximately 15~MeV.

To compare data and simulation, all inoperable or noisy channels in SVT and ECal have been disabled in the simulation, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well. Using a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_simulation}
	\caption{Weighted E/p from Monte Carlo simulation after gain calibration.}
	\label{fig:gains}
}
\end{center}
\end{figure}

These gains can then be used to convert from ADC counts in a channel to the energy deposited into that ECal 
crystal.
The other information needed to find the energy of an incident particle is the sampling fraction---the ratio of 
energy read out from cluster to energy of an incident particle. The conventional sampling fraction, 
the fraction of incident energy that is deposited in the cluster, is approximately 0.9  for the ECal (somewhat less at the 
edges). For our readout, there is additional energy lost because crystals under the threshold are not 
read out. The weighted E/p used in calibration (see Figure \ref{fig:gains}) is an approximate measurement of sampling fraction, but the sampling fraction is energy-dependent because of the effect of readout threshold. 
A full computation of sampling fraction can be done using simulation.

%\subsubsection{Trigger Performance}
As discussed above, the trigger and DAQ integrate pulses differently to measure hit energy. The trigger 
integrates using a time-over-threshold window, and the DAQ readout integrates using a constant window 
(5 samples before and 30 samples after a threshold crossing). For every event, the trigger reports the 
trigger decision as a bit mask (top trigger, bottom trigger, or both) and the time the trigger fires.

We study trigger performance by simulating the trigger for each event and comparing to how events were 
actually triggered.
First, we simulate the FADC readout board to convert from readout hits (constant integration window) to 
trigger hits (time-over-threshold integration). We then simulate the CTP clustering algorithm and the trigger 
decision before we compare the trigger decision and trigger time reported by the simulation to what was 
reported by the real trigger.

To eliminate trigger bias in checking the trigger decision, we use a tag and probe method. To check trigger 
performance in one half of the ECal, we tag events where there was a trigger in the other half, and exactly one 
probe cluster in the ECal half under test. 
We then measure trigger efficiency (proportion of tagged events where there was a trigger) as a function of 
ADC counts and energy of the probe cluster.
These turn-on curves are shown for the top half of the ECal in Fig.~\ref{fig:turnon}.

\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag}
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag_E}
	\caption{Trigger turn-on as a function of probe cluster ADC counts (top) and probe cluster energy in MeV
	(bottom). Both plots are for the top half of the ECal; bottom is similar. Energy is not corrected for sampling 
	fraction.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger threshold is seen to be 1280 ADC counts as expected. The threshold is not perfectly sharp in this 
analysis because of uncertainties in the conversion from readout to trigger hits but based on comparisons 
with Monte Carlo simulation we believe the trigger worked exactly as specified. 
The trigger threshold in terms of cluster energy is very uneven for two reasons; gain variations between 
different ECal crystals lead to threshold variations and the nonlinearity of the time-over-threshold integral 
means that the effective threshold is higher for clusters that span multiple crystals.
Overall the trigger appears to have functioned exactly as intended. Changes planned for the next run 
(constant integration window and per-crystal gain calibration constants for the trigger) will solve both of the 
issues that led to threshold variations in the test run.

@@

\section{Multiple Coulomb Scattering Distributions}

Occupancies close to the beam create many of the key challenges in the HPS experiment
and determine the limits of sensitivity to low \Aprime{} masses.
These occupancies are dominated by electrons which have multiple Coulomb scattered (MCS) 
to relatively large angles in the target. Because HPS is sensitive to scattering angles far out on the tail of the 
MCS distribution, well beyond the angles important in other experiments, care must be taken
to ensure our simulations are correct in this regime.

@@

Data was taken at three different converter thicknesses with a beam current varying between 
$30-70$~nA, see~Tab.~\ref{tab:currents}.  The photon beam line during the test run produced a relatively 
large fraction of pairs 
originating upstream of the converter. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions 
the same. 
The upstream background measured in the ``empty'' converter runs was subtracted 
from the other runs, properly normalized using the measured integrated currents.
\begin{table}
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickness & Duration &  $e^-$ on converter \\
 (\%$X_0$) & (s) & (nC)    \\   
\hline
%\hline
1.6   & 911 &     24385.9     \\ %27 nA
%\hline
0.18   & 2640 &    193508.9  \\  %73 nA
%\hline
0.45  & 2149 &       140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88079.6  \\
\hline
\end{tabular}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
}
\end{center}
\end{table}

@@

>>>>

The photon beam was generated in the interaction of the $5.5$~GeV electrons with a gold radiator of 
$10^{-4}~X_0$, located $\approx~9$~m upstream of the PS. The primary beam and scattered 
electrons are deflected away from detectors by the dipole magnet of the photon tagging system. After 
collimation (6.4~mm diameter), the photon beam passes through the aluminum PS pair converter and 
later the HPS system. 
The converter was located $\approx~77$~cm upstream of the first layer of the silicon vertex tracker. Data was 
taken on different converters (empty, $1.8\times 10^{-3}~X_0$ , $4.5\times 10^{-3}~X_0$ and 
$1.6 \times 10^{-2}~X_0$). 
These measurements were repeated with the reverse field setting of the PS dipole magnet.

@@

During the duration of the test run all SVT modules and APV25 chips were configured to their 
nominal operating points~\cite{Jones:1069892} with all sensors reverse-biased at 180~V.  The 
sensors were operated within a temperature range of  $20-24^\circ$C 
throughout the test run. 
Throughout the duration of the test run, approximately 97\% of the 12,780 SVT channels were found to
be operating normally. The fraction of dead or noisy channels varied from 2.4\% to 4.7\%.  
Most of these were due to misconfigured readout chips (2--4 misconfigured chips out of 100 ), 
a known noisy half-module and a couple of known noisy readout chips.

@@

After a trigger is received, six samples of the corresponding output of the APV25 shaper circuit are digitized. The samples from every channel on a sensor surviving the 
data reduction algorithm (see Sec.~\ref{sec:svt_daq}) are the basis for offline hit reconstruction. 
The six samples of the APV25 shaper output 
are fitted to an ideal $CR-RC$ function to extract the amplitude and the $t_0$ of the hit. 
The typical pulse shape obtained is shown in Fig.~\ref{fig:pulse_shape} also demonstrates that the SVT 
was well timed in to the trigger with the rise of the pulse at the 3rd sampling point.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_110513_samples_L1_top.pdf}
	\caption{The six pedestal subtracted samples from individual channels associated with a hit on a 
	track.}
	\label{fig:pulse_shape}
}
\end{center}
\end{figure}

These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/run1351_mip_small.pdf}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit 
time with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter 
due to clock phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[h]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/t0_resolution_run1351_110513.pdf}
	\caption{
	The cluster time residual for a representative sensor relative to the track time (see text for more 
	details). \color{red}{\bf Fix axis.} }
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is somewhat 
worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate 
with the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

While occupancy was slightly larger than expected, good agreement between data and simulation was 
found after taking into account dead or noisy channels. The hit efficiency was estimated by measuring 
the number of good tracks with a hit close to the extrapolated intersection of a given sensor that was 
excluded from the track fit itself. Tracks which intersect regions with known bad channels or very close to 
in the edge region are excluded in the calculation of the single hit efficiency. The hit efficiency, see 
Fig.~\ref{fig:hit_efficiency}, was measured to be above 98\% and fairly uniform across the SVT.
\begin{figure}[h]
\begin{center}
{\small
    	\includegraphics[width=7cmh]{figures/hit_efficiency_vs_layer.pdf}
        \caption{ The hit reconstruction efficiency as a function of detector layer.}
	\label{fig:hit_efficiency}
}
\end{center}
\end{figure}
The spatial resolution of similar microstrip sensors is well established by test beam data, against which the 
charge deposition model in the simulation is validated.  This resolution can be parameterized as a 
function of the total signal to single-strip noise (S/N) and the crossing angle of tracks through the sensor.  
The single-hit resolution for charged particles with signal to noise ration above 20, as demonstrated here, 
is relatively constant at approximately 6~$\mu$m for tracks that are close to normal to the sensors as in HPS.

@@

For the vertexing 
performance the foremost difference compared to electron beam running is that the target was 
located approximately 67~cm from our nominal target position; giving almost collinear tracks in the detector. 
This degrades the vertex resolution along the 
beam line compared to that expected in an electron beam with tracks from the nominal target position. 
Furthermore, tails of the vertex distributions are impossible to study with the finite data sample obtained 
in the test run. Nevertheless, useful information can still be 
obtained by studying the vertex distributions. Figure~\ref{fig:vtx_pos} shows the distance of closest 
approach of the momentum vectors extrapolated in the 
upstream direction from our analyzing magnet, taking into account the measured fringe field of the PS magnet. 
 \begin{center}
{\small
	 \begin{figure*}[ht]
	% Plot info:
	% TestRun-v6 geometry, twotrkfilter, trigsel, hps-java-1.8-SNAPSHOT, 4hits, chi2<10
	\includegraphics[width=5cm]{figures/h_vtx_fr_x_h_vtx_x_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\includegraphics[width=5cm]{figures/h_vtx_fr_y_h_vtx_y_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\includegraphics[width=5cm]{figures/h_vtx_fr_z_h_vtx_z_trigsel4hit_pair1351_twotrkfilt-v6-paper}
	\caption{ Vertex position represented by the distance of 
	closest approach of the extrapolated momentum vectors upstream of the PS magnet taking into account 
	the measured fringe field. The overall shift from zero in the x-direction is 
	due to a 30~mrad rotation of the SVT with respect to the beam line.}
	\label{fig:vtx_pos}
\end{figure*}
}
\end{center}
%While the tails of the vertex distribution expected in electron beam running is not accessible here the 
%fact that the core is relatively well described provides confidence of the description of the amount of 
%material and the model used for simulating multiple scattering in the target.  
%These are crucial for benchmarking the HPS \Aprime{} physics reach since both the mass and vertex 
%resolution that determine the sensitivity are limited by multiple scattering.

@@

The integrated pulse of the each channel was converted to energy by first 
subtracting the pedestal then applying a gain or conversion factor to convert ADC counts to energy. 
The pedestals was measured in special runs where a 100 samples long window was recorded for each 
trigger in which the FADCs sampled the APD signals every 4~ns.  The pedestals were extracted from the 
part of the window before the actual hit in the calorimeter. Modules above a threshold are clustered 
using a simple algorithm described in Sec.~\ref{sec:trigger}. The ECal was operated with an effective 
crystal readout threshold of 73~MeV. The noise level of the ECal modules were measured to be approximately 15~MeV.

To compare data and simulation, all inoperable or noisy channels in SVT and ECal have been disabled 
in the simulation; any efficiency or bias effects that affect the real data should be reflected 
in the simulation. Furthermore, a gain calibration is performed to equalize the gains in the 
simulation to those observed in data by comparing the ECal cluster energy $E$ to the momentum $p$ of 
matched tracks in the SVT.  The ``weighted $E/p$'' for each crystal is computed, representing the 
average $E/p$ for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$.
Iteratively, gains are adjusted until the weighted $E/p$ is similar in data and simulation. The $E/p$ 
in data and simulation can be seen in Fig.~\ref{fig:gains}.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_MC_egs_0_016x0_June5_2013_Sho_Ecal_presentation.png}
	\includegraphics[width=7cm]{figures/eoverp_data_run1351_June5_2013_Sho_Ecal_presentation.png}
	\caption{The ECal energy over track momentum ratio for simulation (top) and data (bottom).\color{red}{\bf (Update style.)}} 
	\label{fig:gains}
}
\end{center}
\end{figure}
These gains can then be used to convert from the ADC counts reported by the FADC in each channels 
to the energy deposited in the crystals. In order to find the energy of the incident particle one also needs 
to know the  the fraction of incident energy that is deposited in the cluster or "sampling fraction". 
The sampling fraction is parameterized from simulation that show that it is approximately 0.9 for the 
ECal. The weighted $E/p$ shown in Fig.~\ref{fig:gains} is an approximate measurement of the sampling 
fraction of that part of the ECal but is position- (somewhat smaller at the edges) and energy-dependent 
because of the high readout thresholds (explaining the low value compared to 0.9). Since the 
sampling fraction is not needed to normalize event rates it was not applied in the multiple scattering 
analysis in Sec.~\ref{sec:mcs}.

%\subsubsection{Trigger Performance}
The trigger and DAQ integrate pulses from the ECal modules differently to measure the energy 
deposited in the each crystal. The trigger integrates the pulse using a time-over-threshold window, and 
the DAQ readout integrates using a fixed-size window (5 samples before and 30 samples after a 
threshold crossing). For every event, the trigger reports the trigger decision as a bit mask (top half, 
bottom half or both) and the time the trigger fired. The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
First, we simulate the FADC readout board to convert from readout hits (with fixed-size window 
integration) to trigger hits (time-over-threshold integration). Secondly, the CTP clustering 
algorithm and the trigger decision is simulated before we compare the trigger decision and trigger time 
to what was reported by the actual trigger. To eliminate trigger bias, we use a tag and probe method. To 
study the trigger performance in one half of the ECal, we select events which triggered the other half 
and where there was exactly one probe cluster in the ECal half under study. We then measure trigger 
efficiency as the fraction of tagged events that fired the trigger in the probe half as a function of the 
probe cluster energy, shown in Fig.~\ref{fig:turnon}. As a cross-check, the turn-on from the trigger 
threshold was measured to be 1280 in units of ADC counts as expected. The threshold was not 
perfectly sharp because of uncertainties in the conversion from readout to trigger hits described above, 
but based on comparisons with simulation we believe the trigger worked exactly as specified.
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag_E}
	\caption{Trigger turn-on in the top half of the ECal (bottom is similar) for data and simulation as a 
	function of probe cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below one for two reasons.  Gain 
variations between different crystals lead to threshold variations and the nonlinearity of the 
time-over-threshold integral means that the effective threshold is higher for clusters that span multiple 
crystals. The effective trigger threshold is therefore dependent on position and energy of the particle 
as well as cluster multiplicity. Overall the trigger appears to have functioned exactly as intended. 
Changes planned for physics running (constant integration window and per-crystal gain calibration 
constants for the trigger) will solve both of the issues that led to threshold variations in the Test run.

@@

\section{Multiple Coulomb Scattering Distributions}
\label{sec:mcs}
Occupancies close to the beam create many of the key challenges in the HPS experiment
and determine the limits of sensitivity to low \Aprime{} masses.
These occupancies are dominated by electrons which have multiple Coulomb scattered (MCS) 
to relatively large angles in the target. Because HPS is sensitive to scattering angles far out on the tail of the 
MCS distribution, well beyond the angles important in other experiments, care must be taken
to ensure our simulations are correct in this regime.

@@

Data was taken at three different converter thicknesses with a beam current varying between 
$30-70$~nA, see~Tab.~\ref{tab:currents}.  The photon beam line during the test run produced a relatively 
large fraction of pairs 
originating upstream of the converter. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions 
the same. 
The upstream background measured in the ``empty'' converter runs was subtracted 
from the other runs, properly normalized using the measured integrated currents.
\begin{table}
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickness & Duration &  $e^-$ on converter \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
}
\end{center}
\end{table}

@@

\section{Acknowledgements}
